---
output:Restoration effect
Population:
  24 control/resto. for 4_river + 1_period
  input data name:DataR_3.gen in download 
  RGB, 

    
  word_document: default
  html_document: default
  pdf_document: default
  ## Project design: 
A-Restoration: FST >> He/Ho >> Ar >> Ne
B- Stocking: FST >> Structure

A: 1- upload data
   2- grouped data >> fst >> plot fst >> CI 
   3- the number of private alleles
   4- mean allelic richness
   5- (O) heterozygosity, expected heterozygosity (virtualize, list)
   6- DAPC (image, curve)
---

```{r}
install.packages("ggplot2", repos = "http://cran.us.r-project.org")
library("ggplot2") 
install.packages("adegenet", repos = "http://cran.us.r-project.org") 
library("adegenet") 
install.packages("dplyr", repos = "http://cran.us.r-project.org") 
library("dplyr") 
install.packages("poppr", repos = "http://cran.us.r-project.org") 
library("poppr") 
install.packages("hierfstat", repos = "http://cran.us.r-project.org") 
library("hierfstat") 
install.packages("reshape2", repos = "http://cran.us.r-project.org") 
library("reshape2") 
install.packages("scales", repos = "http://cran.us.r-project.org") 
library("scales") 
install.packages("RColorBrewer", repos = "http://cran.us.r-project.org") 
library("RColorBrewer") 
install.packages("diveRsity", repos = "http://cran.us.r-project.org") 
library("RColorBrewer") 
```


#### FST and CI ####
########################## 

#Fst_restoration_re_final_1

```{r}
Brown.Trout = read.genepop('/Users/nooralwash/Downloads/DataR_2.gen', ncode = 3L)

```


#2 print
```{r}
Brown.Trout
```



```{r}
popNames(Brown.Trout) = c("CHB","RHB","CHR","RHR","CHC","RHC","CKB","RKB","CKR","RKR","CKC","RKC","CVB","RVB","CVR","RVR","CVC","RVC", "CSB","RSB","CSR","RSR","CSC","RSC")
```



##2- grouped data >> fst >> plot fst >> CI 

```{r}
Brown.Trout_sub = popsub (Brown.Trout, sublist = c("CHB","RHB","CHR","RHR","CHC","RHC","CKB","RKB","CKR","RKR","CKC","RKC","CVB","RVB","CVR","RVR","CVC","RVC", "CSB","RSB","CSR","RSR","CSC","RSC"))

Brown.Trout_fst = genet.dist(Brown.Trout_sub, method = "WC84") %>% round(digits = 3)
Brown.Trout_fst
```

```{r}
Brown.Trout_sub
```


#Refinal_3
```{r}
lab_order = c("CHB","RHB","CHR","RHR","CHC","RHC","CKB","RKB","CKR","RKR","CKC","RKC","CVB","RVB","CVR","RVR","CVC","RVC", "CSB","RSB","CSR","RSR","CSC","RSC")

fst.mat = as.matrix(Brown.Trout_fst)
fst.mat1 = fst.mat[lab_order, ]
fst.mat2 = fst.mat1[, lab_order]

ind = which(upper.tri(fst.mat2), arr.ind = TRUE)
fst.df = data.frame(Site1 = dimnames(fst.mat2)[[2]][ind[,2]],
                    Site2 = dimnames(fst.mat2)[[1]][ind[,1]],
                    Fst = fst.mat2[ ind ])

fst.df$Site1 = factor(fst.df$Site1, levels = unique(fst.df$Site1))
fst.df$Site2 = factor(fst.df$Site2, levels = unique(fst.df$Site2))

fst.df$Fst[fst.df$Fst < 0] = 0

fst.df %>% str
```


```{r}
fst.label = expression(italic("F")[ST])
mid = max(fst.df$Fst) / 2
ggplot(data = fst.df, aes(x = Site1, y = Site2, fill = Fst))+
  geom_tile(colour = "black")+
  geom_text(aes(label = Fst), color="black", size = 2)+
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = mid, name = fst.label, limits = c(0, max(fst.df$Fst)), breaks = c(0, 0.05, 0.10))+
  scale_x_discrete(expand = c(0,0))+
  scale_y_discrete(expand = c(0,0), position = "right")+
  theme(axis.text.y = element_text( colour = "black", size = 5, face = "bold"),
        axis.text.x = element_text(angle= 90, colour = "black", size = 5, face = "bold"),
        axis.title = element_blank(),
        panel.grid = element_blank(),
        panel.background = element_blank(),
        legend.position = "right",
        legend.title = element_text(size = 3, face = "bold"),
        legend.text = element_text(size = 2)
        )

```
```{r}
# Set the desired resolution (600 DPI)
pdf("Figure_1.pdf", width = 6.6, height = 3.3, pointsize = 12, bg = "white", family = "Helvetica")

# Insert your plot code here (replace 'your_plot' with your actual plot object)
your_plot <- ggplot(data = fst.df, aes(x = Site1, y = Site2, fill = Fst)) +
  geom_tile(colour = "black") +
  geom_text(aes(label = Fst), color = "black", size = 2) +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = mid, name = fst.label, limits = c(0, max(fst.df$Fst)), breaks = c(0, 0.05, 0.10)) +
  scale_x_discrete(expand = c(0, 0)) +
  scale_y_discrete(expand = c(0, 0), position = "right") +
  theme(axis.text.y = element_text(colour = "black", size = 5, face = "bold"),
        axis.text.x = element_text(angle = 90, colour = "black", size = 5, face = "bold"),
        axis.title = element_blank(),
        panel.grid = element_blank(),
        panel.background = element_blank(),
        legend.position = "right",
        legend.title = element_text(size = 3, face = "bold"),
        legend.text = element_text(size = 2))

# Print your plot
print(your_plot)

# Close the PDF device
dev.off()


```


#C.I. for Fst


```{r}
example_fst = diffCalc (infile="/Users/nooralwash/Downloads/DataR_2.gen", fst = T, bs_pairwise = T, boots = 10000, outfile = "example_fst") 
``` 

```{r}
example_fst
```

#convert list to a matrix Fst_CI (still missing)
###############################

#### Allele ####
########################## 




#Print the number of alleles per locus

```{r}
table(Brown.Trout$loc.fac)

```

#Print the sample size for each site
```{r}
summary(Brown.Trout$pop)

```

#Print the number of private alleles per site across all loci 

```{r}
private_alleles(Brown.Trout) %>% apply(MARGIN = 1, FUN = sum)
```

#Print mean allelic richness per site across all loci (a higher number of alleles at a particular locus can indicate greater genetic diversity.)
```{r}
allelic.richness(genind2hierfstat(Brown.Trout))$Ar %>%
  apply(MARGIN = 2, FUN = mean) %>% 
  round(digits = 3)
```

#Calculate heterozygosity per site (a higher proportion of heterozygous individuals at a particular locus can indicate greater genetic diversity.)


```{r}
basic_Btrout = basic.stats(Brown.Trout, diploid = TRUE)
Ho_Btrout = apply(basic_Btrout$Ho, MARGIN = 2, FUN = mean, na.rm = TRUE) %>%  round(digits = 2)
Ho_Btrout
```

# Mean expected heterozygosity per site

```{r}
He_Btrout = apply(basic_Btrout$Hs, MARGIN = 2, FUN = mean, na.rm = TRUE) %>%
  round(digits = 2)
He_Btrout
```

#Visualise heterozygosity per site (the observed heterozygosity (Ho) and the expected heterozygosity (He))

```{r}

Het_Btrout_df = data.frame(Site = names(Ho_Btrout), Ho = Ho_Btrout, He = He_Btrout) %>%
  melt(id.vars = "Site")

custom_theme = theme(
  axis.text.x = element_text(size = 10, angle = 90, vjust = 0.5, face = "bold"),
  axis.text.y = element_text(size = 10),
  axis.title.y = element_text(size = 12),
  axis.title.x = element_blank(),
  axis.line.y = element_line(size = 0.5),
  legend.title = element_blank(),
  legend.text = element_text(size = 12),
  panel.grid = element_blank(),
  panel.background = element_blank(),
  plot.title = element_text(hjust = 0.5, size = 15, face="bold")
  )
hetlab.o = expression(italic("H")[o])
hetlab.e = expression(italic("H")[e])

ggplot(data = Het_Btrout_df, aes(x = Site, y = value, fill = variable))+
  geom_bar(stat = "identity", position = "dodge", colour = "black")+
  scale_y_continuous(expand = c(0,0), limits = c(0,0.750), breaks = c(0, 0.10, 0.20, 0.30, 0.40, 0.50, 0.60))+
  scale_fill_manual(values = c("pink", "#bdbdbd"), labels = c(hetlab.o, hetlab.e))+
  ylab("Heterozygosity")+
  ggtitle("B.trout Heterozygosity")+
  custom_theme
```


```{r}
nl= c("CHB","RHB","CHR","RHR","CHC","RHC","CKB","RKB","CKR","RKR","CKC","RKC","CVB","RVB","CVR","RVR","CVC","RVC", "CSB","RSB","CSR","RSR","CSC","RSC")
```

```{r}
Het_Btrout_df$Site= factor(Het_Btrout_df$Site, levels = nl)

# Filter the data frame to include only "Ho" (Ho_Btrout)
Het_Btrout_df_filtered <- Het_Btrout_df %>%
  filter(variable == "Ho")

# Custom theme (same as before)
custom_theme <- theme(
  axis.text.x = element_text(size = 10, angle = 90, vjust = 0.5, face = "bold"),
  axis.text.y = element_text(size = 10),
  axis.title.y = element_text(size = 12),
  axis.title.x = element_blank(),
  axis.line.y = element_line(size = 0.5),
  legend.title = element_blank(),
  legend.text = element_text(size = 12),
  panel.grid = element_blank(),
  panel.background = element_blank(),
  plot.title = element_text(hjust = 0.5, size = 15, face = "bold")
)
```


```{r}
# Label for the bar
hetlab_o <- expression(italic("H")[o])

# Create a new variable for alternating colors
Het_Btrout_df_filtered$color <- c("blue", "white")

# Create the plot with alternating colors
ggplot(data = Het_Btrout_df_filtered, aes(x = Site, y = value, fill = color)) +
  geom_bar(stat = "identity", position = "dodge", colour = "black", width = 0.9) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 0.750), breaks = c(0,  0.40, 0.45, 0.50, 0.55, 0.60, 0.65)) +
  scale_fill_identity() +  # Use the color variable directly
  ylab("Heterozygosity") +
  ggtitle("B.trout Ho") +
  custom_theme

```

```{r}
Het_Btrout_df$Site= factor(Het_Btrout_df$Site, levels = nl)

# Filter the data frame to include only "Ho" (Ho_Btrout)
Het_Btrout_df_filtered <- Het_Btrout_df %>%
  filter(variable == "He")

# Custom theme (same as before)
custom_theme <- theme(
  axis.text.x = element_text(size = 10, angle = 90, vjust = 0.5, face = "bold"),
  axis.text.y = element_text(size = 10),
  axis.title.y = element_text(size = 12),
  axis.title.x = element_blank(),
  axis.line.y = element_line(size = 0.5),
  legend.title = element_blank(),
  legend.text = element_text(size = 12),
  panel.grid = element_blank(),
  panel.background = element_blank(),
  plot.title = element_text(hjust = 0.5, size = 15, face = "bold")
)

# Label for the bar
hetlab_o <- expression(italic("H")[o])

# Create a new variable for alternating colors
Het_Btrout_df_filtered$color <- c("blue", "blue")

# Create the plot with alternating colors
ggplot(data = Het_Btrout_df_filtered, aes(x = Site, y = value, fill = color)) +
  geom_bar(stat = "identity", position = "dodge", colour = "black", width = 0.9) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 0.750), breaks = c(0, 0.10, 0.20, 0.30, 0.40, 0.50, 0.60)) +
  scale_fill_identity() +  # Use the color variable directly
  ylab("Heterozygosity") +
  ggtitle("B.trout He") +
  custom_theme

```



```{r}
# Assuming you have already created the filtered data frame Het_Btrout_df_filtered

# Custom theme (same as before)
custom_theme <- theme(
  axis.text.x = element_text(size = 10, angle = 90, vjust = 0.5, face = "bold"),
  axis.text.y = element_text(size = 10),
  axis.title.y = element_text(size = 12),
  axis.title.x = element_blank(),
  axis.line.y = element_line(size = 0.5),
  legend.title = element_blank(),
  legend.text = element_text(size = 12),
  panel.grid = element_blank(),
  panel.background = element_blank(),
  plot.title = element_text(hjust = 0.5, size = 15, face = "bold")
)

# Label for the bar
hetlab_o <- expression(italic("H")[o])

# Create the plot with narrower bars and increased width
ggplot(data = Het_Btrout_df_filtered, aes(x = Site, y = value, fill = color)) +
  geom_bar(stat = "identity", position = "dodge", colour = "black", width = 0.9) +  # Adjust the width
  scale_y_continuous(expand = c(0, 0), limits = c(0, 0.750), breaks = c(0, 0.10, 0.20, 0.30, 0.40, 0.50, 0.60)) +
  scale_fill_identity() +  # Use the color variable directly
  ylab("Heterozygosity") +
  ggtitle("B.trout Heterozygosity") +
  custom_theme

```

#BLue Hetrozygosity
```{r}
nl= c("CHB","RHB","CHR","RHR","CHC","RHC","CKB","RKB","CKR","RKR","CKC","RKC","CVB","RVB","CVR","RVR","CVC","RVC", "CSB","RSB","CSR","RSR","CSC","RSC")
```

```{r}
Het_Btrout_df$Site= factor(Het_Btrout_df$Site, levels = nl)
```

```{r}
ggplot(data = Het_Btrout_df , aes(x = Site, y = value, fill = variable))+
  geom_bar(stat = "identity", position = "dodge", colour = "black")+
  scale_y_continuous(expand = c(0,0), limits = c(0,0.750), breaks = c(0, 0.10, 0.20, 0.30, 0.40, 0.50, 0.60))+
  scale_fill_manual(values = c("#0099ee", "#bdbdbd"), labels = c(hetlab.o, hetlab.e))+
  ylab("Heterozygosity")+
  ggtitle("B.trout Heterozygosity")+
  custom_theme
```


#Inbreeding coefficient (FIS) (The FIS value is a measure of the proportion of homozygosity in a population, with higher values indicating a higher level of homozygosity and lower values indicating a lower level of homozygosity)
Calculate mean FIS per site.

```{r}
Het_Btrout_df
```


```{r}
# Perform cross validation to find the optimal number of PCs to retain in DAPC (Discriminant Analysis Machine Learning model cross validation: Building and cross validating a model that can predict future samples belonging to which population)
set.seed(123)
x = tab(Brown.Trout_sub, NA.method = "mean")
crossval = xvalDapc(x, Brown.Trout_sub$pop, result = "groupMean", xval.plot = TRUE)

```

```{r}
# Number of PCs with best stats (lower score = better)
crossval$`Root Mean Squared Error by Number of PCs of PCA`
##        20        40        60        80       100       120       140 
## 0.7072406 0.6354916 0.6501085 0.6206800 0.6247699 0.6145408 0.6140150
crossval$`Number of PCs Achieving Highest Mean Success`
## [1] "120"
crossval$`Number of PCs Achieving Lowest MSE`
## [1] "140"
numPCs = as.numeric(crossval$`Number of PCs Achieving Lowest MSE`)
```



```{r}
# Run a DAPC using site IDs as priors
dapc1 = dapc(Brown.Trout_sub, Brown.Trout_sub$pop, n.pca = numPCs, n.da = 3)

# Analyse how much percent of genetic variance is explained by each axis
percent = dapc1$eig/sum(dapc1$eig)*100
barplot(percent, ylab = "Genetic variance explained by eigenvectors (%)", ylim = c(0,60),
        names.arg = round(percent, 1))
```

```{r}
# Replace missing data with the mean allele frequencies
x = tab(Brown.Trout_sub, NA.method = "mean")

# Perform PCA
pca1 = dudi.pca(x, scannf = FALSE, scale = FALSE, nf = 3)

# Analyse how much percent of genetic variance is explained by each axis
percent = pca1$eig/sum(pca1$eig)*100
barplot(percent, ylab = "Genetic variance explained by eigenvectors (%)", ylim = c(0,12),
        names.arg = round(percent, 1))
```

```{r}
Brown.Trout_sub
```

```{r}
# Create a data.frame containing individual coordinates
ind_coords = as.data.frame(pca1$li)

# Rename columns of dataframe
colnames(ind_coords) = c("Axis1","Axis2","Axis3")

# Add a column containing individuals
ind_coords$Ind = indNames(Brown.Trout_sub)

# Add a column with the site IDs
ind_coords$Site = Brown.Trout_sub$pop

# Calculate centroid (average) position for each population
centroid = aggregate(cbind(Axis1, Axis2, Axis3) ~ Site, data = ind_coords, FUN = mean)

# Add centroid coordinates to ind_coords dataframe
ind_coords = left_join(ind_coords, centroid, by = "Site", suffix = c("",".cen"))

# Define colour palette
cols = brewer.pal(nPop(Brown.Trout_sub), "Set1")

# Custom x and y labels
xlab = paste("Axis 1 (", format(round(percent[1], 1), nsmall=1)," %)", sep="")
ylab = paste("Axis 2 (", format(round(percent[2], 1), nsmall=1)," %)", sep="")

# Custom theme for ggplot2
ggtheme = theme(axis.text.y = element_text(colour="black", size=12),
                axis.text.x = element_text(colour="black", size=12),
                axis.title = element_text(colour="black", size=12),
                panel.border = element_rect(colour="black", fill=NA, size=1),
                panel.background = element_blank(),
                plot.title = element_text(hjust=0.5, size=15) 
)
# Load the 'viridis' package (if not already installed)
# install.packages("viridis")
library(viridis)


# Scatter plot axis 1 vs. 2
ggplot(data = ind_coords, aes(x = Axis1, y = Axis2)) +
  geom_hline(yintercept = 0) +
  geom_vline(xintercept = 0) +
  geom_segment(aes(xend = Axis1.cen, yend = Axis2.cen, colour = Site), show.legend = FALSE) +
  geom_point(aes(fill = Site), shape = 21, size = 3, show.legend = FALSE) +
  geom_label(data = centroid, aes(label = Site, fill = Site), size = 4, show.legend = FALSE) +
  scale_fill_viridis(discrete = TRUE) +  # Use the 'viridis' fill colors
  scale_colour_viridis(discrete = TRUE) +  # Use the 'viridis' line colors
  labs(x = xlab, y = ylab) +
  ggtitle("B.trout PCA") +
  ggtheme


```

```{r}

x = tab(Brown.Trout_sub, NA.method = "mean")
pca1 = dudi.pca(x, scannf = FALSE, scale = TRUE, nf = 3)
percent = pca1$eig/sum(pca1$eig)*100
ind_coords = as.data.frame(pca1$li)
colnames(ind_coords) = c("Axis1","Axis2","Axis3")
ind_coords$Ind = indNames(Brown.Trout_sub)
ind_coords$Site = Brown.Trout_sub$pop
centroid = aggregate(cbind(Axis1, Axis2, Axis3) ~ Site, data = ind_coords, FUN = mean)
ind_coords = left_join(ind_coords, centroid, by = "Site", suffix = c("",".cen"))
cols = c("#ADD8E6", "#7400DF", "#9C00C0", "#B700A1" , "#CC0083", "#DC0066" ,"#EB0048", "#F70026", "#FD2B10", "#F85C34", "#EF7C53", "#E39871", "red", "#BBC9B1", "#97E0D2", "yellow", "#28FEED", "#3BFED4", "#7FFF00", "#45FEA0", "#42FE84" ,"#3AFE67", "#2BFE45" ,"#00FF00")

xlab = paste("Axis 1 (", format(round(percent[1], 1), nsmall=1)," %)", sep="")
ylab = paste("Axis 2 (", format(round(percent[2], 1), nsmall=1)," %)", sep="")
ggplot(data = ind_coords, aes(x = Axis1, y = Axis2))+
  
  geom_hline(yintercept = 0)+
  geom_vline(xintercept = 0)+
  geom_segment(aes(xend = Axis1.cen, yend = Axis2.cen, colour = Site), show.legend = FALSE, alpha = 0.5)+
  geom_point(aes(fill = Site), shape = 21, size = 2, show.legend = FALSE, alpha = 0.5)+
  geom_label(data = centroid, aes(label = Site, fill = Site), size = 1.8, show.legend = FALSE)+
  scale_fill_manual(values = cols)+
  scale_colour_manual(values = cols)+
  labs(x = xlab, y = ylab)+
  ggtitle(" ")+
theme(panel.background = element_rect(fill = "white")) # Add this line


#  +geom_label_repel(data = centroid, aes(label = Site, fill = Site), size = 4, show.legend = FALSE,
 #                  min.segment.length = 0.5, segment.linetype = 3)



```
```{r}
# Define the original colors
original_colors <- c("blue", "red", "cyan", "green")

# Create a color ramp palette with 24 shades
color_ramp <- colorRampPalette(original_colors, space = "Lab")

# Generate 24 shades
shades_24 <- color_ramp(24)

# Display the shades
print(shades_24)

```

```{r}

# Scatter plot axis 1 vs. 2
ggplot(data = ind_coords, aes(x = Axis1, y = Axis2)) +
  geom_hline(yintercept = 0) +
  geom_vline(xintercept = 0) +
  geom_segment(aes(xend = Axis1.cen, yend = Axis2.cen, colour = Site), show.legend = FALSE) +
  geom_point(aes(fill = Site), shape = 21, size = 2, show.legend = FALSE) +
  geom_label(data = centroid, aes(label = Site, fill = Site), size = 2.5, show.legend = FALSE) +
  scale_fill_viridis(discrete = TRUE) +  # Use the 'viridis' fill colors
  scale_colour_viridis(discrete = TRUE) +  # Use the 'viridis' line colors
  labs(x = xlab, y = ylab) +
  ggtitle("B.trout PCA") +
  ggtheme  +
  theme(plot.margin = margin(0.01, 0.01, 0.01, 0.01, "pt"))  # Adjust margins for larger plot

```

```{r}
# Create a function for jittering
rand_jitter <- function(arr) {
  stdev <- 0.01 * (max(arr) - min(arr))
  return(arr + rnorm(length(arr)) * stdev)
}

# Create a data.frame containing individual coordinates
ind_coords = as.data.frame(dapc1$ind.coord)

# Rename columns of dataframe
colnames(ind_coords) = c("Axis1","Axis2","Axis3")

# Add a column containing individuals
ind_coords$Ind = indNames(Brown.Trout_sub)

# Add a column with the site IDs
ind_coords$Site = Brown.Trout_sub$pop

# Calculate centroid (average) position for each population
centroid = aggregate(cbind(Axis1, Axis2, Axis3) ~ Site, data = ind_coords, FUN = mean)

# Add centroid coordinates to ind_coords dataframe
ind_coords = left_join(ind_coords, centroid, by = "Site", suffix = c("",".cen"))

# Define colour palette
cols = brewer.pal(nPop(Brown.Trout_sub), "Set2")

# Custom x and y labels
xlab = paste("Axis 1 (", format(round(percent[1], 1), nsmall=1)," %)", sep="")
ylab = paste("Axis 2 (", format(round(percent[2], 1), nsmall=1)," %)", sep="")



# Custom theme for ggplot2
ggtheme = theme(axis.text.y = element_text(colour="black", size=12),
                axis.text.x = element_text(colour="black", size=12),
                axis.title = element_text(colour="black", size=12),
                panel.border = element_rect(colour="black", fill=NA, size=1),
                panel.background = element_blank(),
                plot.title = element_text(hjust=0.5, size=15) 
)
```



